---
title: "Network Notes"
author: "Rob Hayward"
date: "26 January 2016"
output: html_document
---

#Introduction
Networks are everywhere.  There are connections called *edges* between *nodes* or *vertices*.  The edges can be *directed* or not. There are two main packages: `igraph`, the work of Gabor Casrdi, and `statnet` package from the University of Washington. 

##Representation
Matrices can be used to represent networks
```{r} 
A <- rbind(c(0,1,0), c(1,0,1), c(2,0,0))
nodeNames <- c("A", "B", "C")
dimnames(A) <- list(nodeNames, nodeNames)
A
```
Where the row are from a node and the columns are to a node. 

An alternative is to use an *edge list* representation. 

However, the `igraph` package has its own efficient internal representation. This is most easily used by using the function `graph_from_literal`

```{r, message=FALSE, warning=FALSE}
library(igraph)
g <- graph_from_literal(A++B, B-+C, C-+A)
g
```
It has efficient internal stricture.  The output indicates that it is directed (D), it is named (N) and that it has 3 nodes and 4 edges. 
```{r, warning=FALSE, message=FALSE,fig.height=4.6}
plot(g)
```

igraph objects can also be made with 

* edge lists `g <- graph_from_edgelist(el, directed = TRUE)`
* data frames `g <- graph_from_data_frame(df, directed = TRUE)`
* adjecency matrix `g <- graph_from_adjacency_matrix(A)`

## From graph to matrix form
It is possible move back from the graph to the matrix form 

###Adjacency matrix

```{r, warning=FALSE, message=FALSE, fig.height=4.6}
A <- as_adjacency_matrix(g, sparse = FALSE)
A
```
###Edge list

```{r, warning=FALSE, message=FALSE, fig.height=4.6}
el <- as_edgelist(g)
el
```

##Standard structures
There are a range of in-built structures 

###Trees

```{r, fig.height=4.6}
g <- make_tree(27, children = 3)
plot(g)
```

###Cliques

```{r, fig.height=4.6}
g <- make_full_graph(n = 6)
plot(g)
```

###Lattices 

```{r, fig.height=4.6}
g <- make_lattice(dimvector = c(5, 5), circular = FALSE)
plot(g)
```

###Stars

```{r, fig.height=4.6}
g <- make_star(n = 10, mode = "undirected")
plot(g)
```

To create a start with 5 nodes and then convet the network to an adjency network

```{r}
g <- make_star(n = 5)
a <- as_adjacency_matrix(g)
a
```